---
title: "Auto Upwelling 3 1980-2020"
author: "Eli Holmes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auto Upwelling 3 1980-2020}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.width = "70%", fig.align = "center")
```

This example shows how work through the points along the 20km coastline and return the upwelling for each day in the year. See `vignette("Auto Upwelling Detection 3", package="imageryML")` for how to do this for one year. This vignette is just the code.

Load the needed packages for plotting and data.
```{r message=FALSE, warning=FALSE}
require(raster)
require(imageryML) # the data
require(magrittr) # for %>%
```

# Download the data

## Download from ERDDAP

Set region lat-lon of interest.
```{r}
min_lat <- 40.625
max_lat <- 50.625
min_lon <- 229.875
max_lon <- 236.625
```

If the file does not exist, then download.
```{r eval=FALSE}
for (yr in 1982:2020) {
  start_date <- paste0(yr, "-01-01")
  stop_date <- paste0(yr, "-12-31")
  file_name <- file.path(here::here(), "inst", "extdata", paste0(
    "erddap_", start_date, "_", stop_date, "_", "lat_", min_lat, "_", max_lat,
    "_lon_", min_lon, "_", max_lon, ".csv"
  ))
  if (file.exists(file_name)) next
  cat(yr, " ")
  import_erddap(start_date, stop_date,
    min_lat = min_lat, max_lat = max_lat,
    min_lon = min_lon, max_lon = max_lon,
    silent = TRUE
  )
}
```

## Function to process raw data frame

## Process data

Notice that I substract 360 off the longitudes. **This is crucial.** The standard projections use this location for the 0 longitude line.

```{r}
processDF <- function(df_raw) {
  df_processed <- df_raw %>%
    # Get rid of miscellaneous zlev in first row
    dplyr::slice(-1) %>%
    # zlev is a column of zeroes, so get rid of that
    dplyr::select(-zlev) %>%
    # Convert into date
    dplyr::mutate(
      time = lubridate::ymd_hms(time),
      # Extract out day so I can just filter for first day in each month
      day = lubridate::day(time)
    ) %>%
    dplyr::select(-day) %>%
    # Set column names
    dplyr::rename(
      date = time,
      lat = latitude,
      lon = longitude
    ) %>%
    # Convert date column to Date type
    dplyr::mutate(
      date = as.Date(date),
      lat = as.numeric(lat),
      lon = as.numeric(lon) - 360,
      sst = as.numeric(sst)
    )
  return(df_processed)
}
```

## Set up needed lines and raster

Need to create a sample raster on the Winkel Tripel projection (what the buffer 300km line is in) and then get its bounding box and crop our 300km line to that.
```{r message=FALSE}
efil <- "erddap_2010-01-01_2010-12-31_lat_40.625_50.625_lon_229.875_236.625.csv"
filename = system.file("extdata", efil, package = "imageryML")
df_raw <- readr::read_csv(filename)
df_processed <- processDF(df_raw)
custom_date <- "2010-08-04"
x <- df_processed %>% dplyr::filter(date == custom_date)
r <- x %>% dplyr::select(lon, lat, sst) %>% raster::rasterFromXYZ(crs="+proj=longlat")

# Re-project the SST raster into Winkel Tripel projection (which is what 300km line is in)
wintri.crs <- raster::crs(sample_points$km100, asText = TRUE)
r.wintri <- raster::projectRaster(r, crs = wintri.crs, over = TRUE)

# Crop down the 300km line
r.wintri.bbox <- raster::bbox(r.wintri)
l.offshore <- smoothr::smooth(buffer300$wintri$line, method = "ksmooth", smoothness = 5)
l.offshore <- raster::crop(l.offshore, r.wintri.bbox)
```

## Run through the years

```{r eval=FALSE}
final_df <- c()
for (yr in 1982:2020) {
  cat(yr, " ")
  start_date <- paste0(yr, "-01-01")
  stop_date <- paste0(yr, "-12-31")
  efil <- paste0("erddap_", start_date, "_", stop_date, "_", "lat_", min_lat, "_", max_lat, "_lon_", min_lon, "_", max_lon, ".csv")
  filename <- system.file("extdata", efil, package = "imageryML")
  df_raw <- readr::read_csv(filename, show_col_types = FALSE)
  df_processed <- processDF(df_raw)
  dates <- unique(df_processed$date)
  for (i in 1:length(dates)) {
    x <- df_processed %>% dplyr::filter(date == dates[i])
    out <- autoDetect3(x, l.offshore = l.offshore, threshold = 2, val = "sst", smooth.method = "none")$df
    out <- cbind(date = dates[i], out)
    final_df <- rbind(final_df, out)
  }
}
final_df <- as.data.frame(final_df)
```

## Add on the lat-lon values

For making plots, it will be helpful to have the lat-lon values for the coast points.
```{r eval=FALSE}
# 1. make the x and y into spatial points. need to use the wintri projection
tmp <- sp::SpatialPoints(final_df %>% dplyr::select(x, y), proj4string = crs(wintri.crs))
# 2. transform into the +proj=longlat; I am using crs(r) just to make sure I use the same crs as the original raster (which should be "+proj=longlat")
tmp <- sp::spTransform(tmp, crs(r))
# 3. Add the lat-lon onto final_df
final_df_wlatlon <-cbind(final_df, lon=tmp@coords[,1], lat=tmp@coords[,2])
final_df_wlatlon <- final_df_wlatlon[c("date", "x", "y", "lon", "lat", 
                       "offshore.SST", "coast.SST", "upwelling")]
```

```{r include=FALSE, eval=FALSE}
file_name <- file.path(here::here(), "inst", "extdata", "final_df_wlatlon.rda")
save(final_df_wlatlon, file=file_name)
```

## Summary

`final_df_wlatlon` now has the upwelling logical value for each day from 1982 to 2020. This data frame is save in `extdata` as `final_df_wlatlon.rda`. This data.frame can be used for making heatmaps and contour maps of the upwelling timing.
