---
title: "Smooth Coastline Data"
author: "Eli Holmes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Smooth Coastline Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
knitr::opts_chunk$set(out.width = "70%", fig.align = "center")
```

This example shows how the data in the package were created and saved. The details for how to do these calculations is in the articles in the [basics](https://ocean-satellite-tools.github.io/basics/) package in my ocean satellite tools org.

Specify where to save the downloaded data. Data from a `rerddap::griddap()` call can be verbose, so you'll probably want a folder for it.
```{r create_dir}
fil_dir <- file.path(here::here(), "inst", "extdata")
if(!dir.exists(fil_dir)) dir.create(fil_dir)
```

Winkel Tripel projection
```{r}
crs.wintri <- "+proj=wintri +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=km +no_defs"
```


## `sample_raster`

```{r}
lats <- c(40.375, 50.375)
lons <- c(-141.875, -120.875)
df_info <- rerddap::info("ncdcOisst21Agg_LonPM180")
df <- rerddap::griddap("ncdcOisst21Agg_LonPM180", latitude = lats, longitude = lons, time = c("2021-06-19", "2021-06-19"), fields = "sst")$data
```

We turn this matrix into a raster using the **raster** package and the `rasterFromXYZ()` function. 
```{r}
df2 <- data.frame(x=df$lon, y=df$lat, z=df$sst)
ras <- raster::rasterFromXYZ(df2, crs = "+proj=longlat")
```

Convert to Winkel Tripel projection
```{r}
ras.wintri <- raster::projectRaster(ras, crs=crs.wintri, over=TRUE)
lats.wintri <- raster::bbox(ras.wintri)[1,]
lons.wintri <- raster::bbox(ras.wintri)[2,]
```

```{r}
data_dir <- file.path(here::here(), "data", "sample_raster.rda")
sample_raster <- list(df=df, raster=ras, lats=lats, lons=lons, 
                      raster.wintri=ras.wintri, lats.wintri=lats.wintri, 
                      lons.wintri=lons.wintri, crs=crs.wintri)
save(sample_raster, file=data_dir)
```

## World coastline

```{r}
world <- rnaturalearth::ne_countries(scale = "small", returnclass = "sp")
# Get rid of interior boundaries
world <- rgeos::gUnaryUnion(world)
```

```{r}
world.wintri <- sp::spTransform(world, crs.wintri)
```

# This part goes through and only saves big polygons to get rid of little islands
```{r}
p <- lapply(world.wintri@polygons , slot , "Polygons")
big <- log(unlist(lapply(p[[1]], function(x) methods::slot(x, "area")))) > 14
# Only keep the continents
trim.world.wintri <- world.wintri
trim.world.wintri@polygons[[1]]@Polygons <- trim.world.wintri@polygons[[1]]@Polygons[which(big)]
trim.world.wintri@polygons[[1]]@plotOrder <- 1:length(trim.world.wintri@polygons[[1]]@Polygons)
slot(trim.world.wintri, "polygons") <- lapply(methods::slot(trim.world.wintri, "polygons"),
   "comment<-", NULL)
```

Here is a plot where the smaller polygons (grey) have been trimmed.
```{r eval=TRUE, message=FALSE}
library(imageryML)
require(raster)
plot(world.wintri, col="grey")
plot(trim.world.wintri, col="green", add=TRUE)
```

## 300 km offshore line

Note, the units for `world.wintri` were set to km. It is important that this code is run on 
a polygon in a meter projection as opposed to a long-lat projection.

The first polygon in the `world` object is the North-South American polygon. I am only going to use that one here.
```{r}
buffer300 <- list()
polys <- list(wintri=world.wintri, trim.wintri=trim.world.wintri)
for(j in 1:length(polys)){
world.poly <- polys[[j]]
buff1 <- rgeos::gBuffer(world.poly, width = 300, byid = TRUE)
# erase the inner world polygon
e <- raster::erase(buff1, world.poly)
# Use the `remove.holes()` function to get only the outer line.
e300 <- spatialEco::remove.holes(spatialEco::remove.holes(e))
# Only keep the continents
e300@polygons[[1]]@Polygons <- e300@polygons[[1]]@Polygons[2]
e300@polygons[[1]]@plotOrder <- as.integer(1)
el300 <- methods::as(e300, "SpatialLines")
df <- c()
n <- length(el300@lines[[1]]@Lines)
for (i in 1:n) {
  df <- rbind(df, cbind(el300@lines[[1]]@Lines[[i]]@coords, ID = i))
}
tmp <- list(line = el300, polygon = e300, df = df, crs = crs.wintri)
buffer300[[names(polys)[j]]] <- tmp
}
```

Smooth the 300 km buffer to get rid of wiggles that will affect nearest point calculations.
```{r}
el <- smoothr::smooth(el300, method="ksmooth", smoothness=5)
df <- c()
n <- length(el@lines[[1]]@Lines)
for (i in 1:n) {
  df <- rbind(df, cbind(el@lines[[1]]@Lines[[i]]@coords, ID = i))
}
tmp <- list(line = el, df = df, crs = crs.wintri)
buffer300[["wintri.smooth"]] <- tmp
```

```{r eval=TRUE, fig.cap="*Red line is the 300km offshore line.*"}
plot(trim.world.wintri, col="grey")
plot(buffer300[["wintri.smooth"]][["line"]], add = TRUE, col="red")
```

## Make 20km coast line

Only Holes function direct adaptation from `spatialEco::remove.holes()`

```{r}
only.holes <- function (x) 
{
  if (!any(which(utils::installed.packages()[, 1] %in% "maptools"))) 
    stop("please install maptools package before running this function")
  xp <- slot(x, "polygons")
  holes <- lapply(xp, function(x) sapply(methods::slot(x, "Polygons"), 
                                         methods::slot, "hole"))
  res <- lapply(1:length(xp), function(i) methods::slot(xp[[i]], 
                                                        "Polygons")[holes[[i]]])
  IDs <- row.names(x)
  x.fill <- sp::SpatialPolygons(lapply(1:length(res), function(i) sp::Polygons(res[[i]], 
                                                                               ID = IDs[i])), proj4string = sp::CRS(sp::proj4string(x)))
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), maptools::checkPolygonsHoles)
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), "comment<-", NULL)
  pids <- sapply(methods::slot(x.fill, "polygons"), function(x) methods::slot(x, "ID"))
  x.fill <- sp::SpatialPolygonsDataFrame(x.fill, data.frame(row.names = pids, 
                                                            ID = 1:length(pids)))
  return(x.fill)
}
```

Create buffer 280 km around the 300km line. Note because we are doing this around a line, we don't need to erase the inner polygon as we had to do for the 300km line. Again this is only for the North-South America polygon.
```{r}
buffer20 <- list()
polys <- list(wintri=buffer300$wintri$line, trim.wintri=buffer300$trim.wintri$line)
for(j in 1:length(polys)){
  p <- polys[[j]] # really line not poly
buff20 <- rgeos::gBuffer(p, width=280)
# Remove the outer part of the polygons and make into lines.
e20 <- only.holes(buff20)
el20 <- methods::as(e20, "SpatialLines")
# Convert to a dataframe also.
df <- c()
n <- length(el20@lines[[1]]@Lines)
for (i in 1:n) {
  df <- rbind(df, cbind(el20@lines[[1]]@Lines[[i]]@coords, ID = i))
}
tmp <- list(line = el20, polygon = e20, df = df, crs = crs.wintri)
buffer20[[names(polys)[j]]] <- tmp
}
```

```{r eval=TRUE, fig.cap="*Red line is the 300km offshore line. Blue line is the 20km offshore line.*"}
plot(trim.world.wintri, col="grey")
plot(buffer300[["trim.wintri"]][["line"]], add = TRUE, col="red")
plot(buffer20[["trim.wintri"]][["line"]], add = TRUE, col="blue")
```

We can plot the "trim" lines (islands removed) and see that the trim line will run through islands. Depending on the goal, that may or may not be a problem.
```{r eval=TRUE}
plot(sample_raster$raster.wintri)
plot(world.wintri, col="grey", add=TRUE)
plot(sample_raster$raster.wintri, add=TRUE, alpha=0.75)
plot(buffer20[["trim.wintri"]][["line"]], add=TRUE, col="red")
plot(buffer20[["wintri"]][["line"]], add=TRUE, col="blue")
```

## Coast sample locations every 100km

Next we want to sample the points every 100km along the coast.

```{r}
sample_points <- list()
polys <- list(wintri=buffer300$wintri$line, trim.wintri=buffer300$trim.wintri$line)
for(j in c("wintri", "trim.wintri")){
  el20 <- buffer20[[j]][["line"]]
  numOfPoints  <-  rgeos::gLength(el20) / 100
  sample.pts.100km <- sp::spsample(el20, n = numOfPoints, type = "regular")
  el300 <- buffer300[[j]][["line"]]
  numOfPoints  <-  rgeos::gLength(el300) / 100
  sample.pts.off.100km <- sp::spsample(el300, n = numOfPoints, type = "regular")
  tmp <- list(km100=sample.pts.100km, off.km100=sample.pts.off.100km)
  sample_points[[j]] <- tmp
}
```

We will not use the trim line since we don't want our samples to run into the islands.
```{r eval=TRUE}
plot(sample_raster$raster.wintri)
plot(world.wintri, col="grey", add=TRUE)
plot(sample_raster$raster.wintri, add=TRUE, alpha=0.75)
plot(buffer300[["wintri"]][["line"]], add=TRUE)
plot(buffer20[["wintri"]][["line"]], add=TRUE, col="red")
plot(sample_points[["wintri"]]$km100, add=TRUE, pch=1)
plot(sample_points[["wintri"]]$off.km100, add=TRUE)
```

## Save the objects

```{r}
crs_wintri <- crs.wintri
save(crs_wintri, file = file.path(here::here(), "data/crs_wintri.rda"))
save(world, world.wintri, trim.world.wintri, file = file.path(here::here(), "data/world.rda"))
save(buffer300, buffer20, file = file.path(here::here(), "data/buffers.rda"))
save(sample_points, file = file.path(here::here(), "data/sample_points.rda"))
```

